### 2018-05-06 paper/main.tex
Paper progress
Theory section is mostly done. This was the hardest part almost certainly.

\documentclass{article}

\usepackage[nonatbib,preprint]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{optidef}
\usepackage{stmaryrd}
\usepackage{mathtools}
\usepackage{algpseudocode}
\usepackage{algorithm}

\usepackage[
  backend=biber,
  style=nature,
  sorting=ynt
]{biblatex}
\addbibresource{main.bib}

\newcommand{\RR}{\mathbb{R}}
\DeclarePairedDelimiter{\norm}{\|}{\|}
\DeclarePairedDelimiter{\ip}{\langle}{\rangle}
\DeclarePairedDelimiter{\ex}{\mathbb{E}[}{]}
\newcommand{\inv}{{}^{-1}}
\newcommand{\trans}{{}^{\top}}

\title{A Stochastic Quasi-Newton Optimizer for TensorFlow}
\author{
  Jason Chen \\
  Department of Computer Science \\
  Stony Brook University \\
  Stony Brook, NY 11790 \\
  \texttt{hungrchen@cs.stonybrook.edu} \\
  \And
  David Kraemer \\
  Department of Applied Mathematics\\
  Stony Brook University \\
  Stony Brook, NY 11790 \\
  \texttt{davidkraemer@stonybrook.edu}
}

\date{May 6, 2018}

\begin{document}

\maketitle

\begin{abstract}
  We do some shit and it doesn't work.
\end{abstract}

\section{Introduction}

\begin{enumerate}
  \item Introduce problem that SdLBFGS solves.
  \item TensorFlow current state of the art.
  \item Problems with the TensorFlow state.
  \item Overview of paper.
\end{enumerate}

\section{Review of theory}

\begin{enumerate}
  \item Review main results of the Wang, Ma, Goldfarb, Liu paper
    \cite{sdlbfgs}
  \item State the SdLBFGS algortihm.
  \item Mention some details about how the algorithm translates to code.
\end{enumerate}

The SdLBFGS algorithm is proposed by Wang, Ma, Goldfarb, and Liu \cite{sdlbfgs}
for solving nonconvex stochastic optimization problems. In particular, it solves
\begin{mini}
  {x \in \RR^n}{f(x) = \ex{F(x, \xi)}}{}{} 
  \label{prb:mini}
\end{mini}
for $F \in C^1(\RR^n \times \RR^d)$ and $\xi \in \RR^d$ is a random variable
with cumulative distribution function $P$. Here $\ex{\cdot}$ denotes the
mathematical expectation. In general $F$ need not be convex.

The function $F$ is readily interpreted in the context of supervised statistical
learning.  Given a corpus $\Xi$ of data, we put a distribution on $\Xi$ which
reflects a sampling regime for a training set. The object of the learning
process is, then, to the model parameter $x \in \RR^n$ which minimizes $F$
\emph{in expectation} with respect to the training set. Provided that the
sampling maintains some resemblance to the overall corpus, this gives a useful
approximate parameter for the whole corpus.

The SdLBFGS algorithm emerges from two concurrent optimization techniques.
First, it follows in the tradition of approximate second-order algorithms which
seek to perform a version of Newton's method while avoiding the computation of
the Hessian $\nabla^2 f(x)$ explicitly. The classical BFGS approach is the
canoncal quasi-Newton method. Second, it follows the development of stochastic
algorithms which seek to randomly sample the components of the gradient $\nabla
f(x)$ so that the essential convergence properties hold in expectation. In this
sense it succeeds Stochastic Gradient Descent, among other canonical methods of
this class. The limited memory constraint has antecedents in both streams of
optimization techniques.

Wang et al. \cite{sdlbfgs} assume that Problem (\ref{prb:mini}) satisfies the
following assumptions:
\begin{enumerate}
  \item There is a real lower bound for $f$. That is, $\inf_{x \in \RR^n} f(x) >
    -\infty$.
  \item The gradient $\nabla f$ is Lipschitz continuous. That is, there exists
    $L > 0$ such that 
    \begin{equation*}
      \norm{\nabla f(x) - \nabla f(y)} \leq L \norm{x - y}
    \end{equation*}
    for all $x,y \in \RR^n$. Moreover, $\nabla_{xx} F(x, \xi)$ exists and is
    continuous and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
    \varepsilon)} \leq \kappa$ for any $x, \xi$.
  \item The sampling regime is unbiased. That is, for the $k$th iterate the
    following hold:
    \begin{align*}
      \ex{g(x_k, \xi_k)}_{\xi_k} &= \nabla f(x_k), \\
      \ex{\norm{g(x_k, \xi_k) - \nabla f(x_k)}^2}_{\xi_k} &\leq \sigma^2,
    \end{align*}
    where $g(x, \xi_k) = \nabla F(x, \xi_k)$ and $\sigma^2$ is the noise level
    of estimating $g(x, \xi_k)$. 
\end{enumerate}

For a convex function, $\nabla^2 f(x)$ is alwayas positive definite, but this
need not be the case in for nonconvex functions. The work in Wang et al.
\cite{sdlbfgs} shows how the BFGS approximate Hessian may be tweaked to ensure
that positive definiteness is preserved even in nonconvex cases. The overall
stochastic quasi-Newton method is described in Algorithm \ref{alg:sqn}, while
the detailed implementation of SdLBFGS is given in Algorithm \ref{alg:sdlbfgs}.

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SQN}{$x_1, \ip{m_k}_{k=1}^{\infty},
    \ip{\alpha_k}_{k=1}^{\infty}, p$}
    \State{$\textrm{data} \leftarrow \emptyset$}
    \For {$k = 1, 2, \dots$}
    \State{$g_k \leftarrow \frac{1}{m_k} \sum_{i=1}^{m_k} g(x_k,
    \varepsilon_{k,i})$} 
    \State{$\Delta x_k, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1} \leftarrow
    \textsc{SdLBFGS}(g_k, {}^*(\textrm{data}\trans))$} \Comment{Tuple unpacking}
    \State{$x_{k+1} \leftarrow x_k - \alpha_k \Delta x_k$}
    \State{$\textit{append}(\textrm{data}, (s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}))$}
    \If{$k > p$}
    \State{$\textit{pop}(\textrm{data})$}
    \EndIf
    \EndFor
    \EndFunction
  \end{algorithmic}
  \label{alg:sqn}
  \caption{%
    The high level stochastic quasi-Newton minimization algorithm. Given an
    initial point $x_1 \in \RR^n$, batch sizes $\ip{m_k}$, step sizes
    $\ip{\alpha_k}$, and a maximum memory capacity $p$, perform BFGS with the
    stochastic dampened direction update.
  }
\end{algorithm}

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SdLBFGS}{$g_{k-1}, \ip{s_{j}}_{j=k-p}^{k-2},
    \ip{\bar{y}_{j}}_{j=k-p}^{k-2}, \ip{\rho_{j}}_{j=k-p}^{k-2}$}
    \State{$\textrm{mem} \leftarrow \textit{min}(p, k-1)$}
    \State{$s_{k-1} \leftarrow x_k - x_{k-1}$}
    \State{$y_{k-1} \leftarrow \frac{1}{m_{k-1}}\sum_{i=1}^{m_{k-1}}[g(x_k,
    \xi_{k-1,i}) - g(x_{k-1}, \xi_{k-1,i})]$}
    \State{$\gamma_k \leftarrow \textit{max}\left( \frac{y_{k-1}\trans
    y_{k-1}}{s\trans_{k-1}y_{k-1}}, \delta \right) \geq \delta$}
    \State{%
      $\theta \leftarrow
      \textit{max}\left( 
        \frac{3}{4} \frac{\gamma_{k}\inv s_{k-1}\trans
        s_{k-1}}{\gamma_k\inv s_{k-1}\trans s_{k-1} - s_{k-1}\trans y_{k-1}}, 1
      \right)
    $}
    \Comment{$\theta$ preserves positive definiteness.}
    \State{$\bar{y}_{k-1} \leftarrow \theta y_{k-1} + (1 -\theta \gamma_{k}\inv s_{k-1})$}
    \State{$\rho_{k-1} \leftarrow (s_{k-1}\trans \bar{y}_{k-1})\inv$}
    \For{$i =0 , \dots, \textrm{mem}-1$}
    \State{$\mu_i \leftarrow \rho_{k-i-1} u_i\trans s_{k-i-1}$} 
    \State{$u_{i+1} \leftarrow u_i - \mu_i \bar{y}_{k-i-1}$}
    \EndFor
    \State{$v_0 \leftarrow \gamma_k\inv u_{\textrm{mem}}$}
    \For{$i =0 , \dots, \textrm{mem}-1$}
    \State{$\nu_i \leftarrow \rho_{k-\textrm{mem}+i} v_i\trans
    \bar{y}_{k-\textrm{mem}+i}$}
    \State{$v_{i+1} \leftarrow v_i + (\mu_{\textrm{mem}-i-1} - \nu_i)s_{k-\textrm{mem}+i} $}
    \EndFor
    \State{\Return{$v_p, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}$}}
    \EndFunction
  \end{algorithmic}
  \caption{%
    The SdLBFGS update step. The resulting output is $H_k g_k = v_p$,
    where $H_k$ is the approximation of the $k$th iterate Hessian and $g_k$ is the
    approximation of the $k$th iterate gradient. Note that the computation of $H_k$
    is implicit, preventing additional storage requirements.
  }
  \label{alg:sdlbfgs}
\end{algorithm}

\section{Implementation}

\begin{enumerate}
  \item Describe how TensorFlow algorithms are usually implemented.
  \item Contrast with our implementation of SdLBFGS.
  \item Explain some implementation issues.
\end{enumerate}

\section{Empirical results}

\begin{enumerate}
  \item Results on simple optimization problems. 
  \item Results on potentially more difficult optimization problems.
  \item Lack of results on MNIST with TensorFlow.
\end{enumerate}

\section{Discussion}

\begin{enumerate}
  \item Explaining why the basic project --- implementing a useful practical
    TensorFlow optimzier --- failed.
  \item Takeaways from the results.
  \item Future work.
\end{enumerate}

\printbibliography[heading=bibintoc]

\end{document}